---
title: "cma-week2_exercice"
author: "Martine Besse"
date: "2023-04-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1: import data

```{r}
library("readr") 
library("sf")
library("ggplot2")
library("dplyr")
```


```{r}
#import csv
wildschwein_BE <- read_delim("data/wildschwein_BE_2056.csv", ",")

#convert to spatial object
wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056, remove = FALSE) #remove = False so the geometry column stays 
```


## Task 2: Getting an overview
Calculate the time difference:  
```{r}
mutate(wildschwein_BE, timelag = as.numeric(difftime(lead(DatetimeUTC), DatetimeUTC))) %>% View()
```

```{r}
#arrange the data
arrange(wildschwein_BE, TierName, DatetimeUTC)
```

How many individuals were tracked?  
`r wildschwein_BE$TierName %>% n_distinct() `
`r wildschwein_BE$TierName %>% unique() `
```{r}
wildschwein_BE %>% group_by(TierName) %>% summarise(n_observations=n())
```

For how long were the individual tracked? Are there gaps? 
Were all individuals tracked concurrently or sequentially?  
What is the temporal sampling interval between the locations?  
```{r}
ggplot(wildschwein_BE, aes(DatetimeUTC, TierName)) +
  geom_point()
```

```{r}
wildschwein_BE <- wildschwein_BE %>%
  group_by(TierName) %>%
  mutate(diff_s = as.numeric(difftime(lead(DatetimeUTC), DatetimeUTC)))
```

```{r}
ggplot(wildschwein_BE, aes(diff_s/60)) +
  geom_histogram(bindwidth = 1) +
  lims(x = c(0,5000/60)) +
  scale_y_log10()
```

```{r}
ggplot(wildschwein_BE, aes(DatetimeUTC, diff_s, colour = TierName)) +
  geom_point()
```

Finding the sampling by changing the datetime:  

```{r}
wildschwein_BE %>%
  filter(DatetimeUTC < "2014-08-24") %>%
  ggplot(aes(DatetimeUTC, diff_s, colour = TierName)) +
  geom_point() +
  geom_line()
```

With 2014-08-24, we can see the difference between day and night pattern (wildboars sleep during the day)


## Task 3:  Deriving movement parameters I: Speed

Try the formula first:  
```{r}
# current location 1 and consecutive location 2
N1 <- 1204752
N2 <- 1204863
E1 <- 2570409
E2 <- 2570402

#formula
#distance from first location to second location
sqrt((E1-E2)^2 + (N1-N2)^2)
```

Implement function in our dataframe with mutate
```{r}
wildschwein_BE <- wildschwein_BE %>%
  group_by(TierName) %>%
  mutate(steplength_m = sqrt((E-lead(E))^2 + (N-lead(N))^2))
  
wildschwein_BE %>% View()
```

Speed:
```{r}
wildschwein_BE <- wildschwein_BE %>%
  mutate(speed_ms = steplength_m/diff_s)
```

```{r}
hist(log10(wildschwein_BE$speed_ms), 100)
```

## Task 4: Cross-scale movement analysis
```{r}
#import csv
caro <- read_delim("data/caro60.csv", ",")
```


Now manually reduce the granularity of our sampling interval by selecting every 3rd, 6th and 9th position and save the output to caro_3, caro_6,caro_9 accordingly.
```{r}
seq_3 <- seq(from = 1, to = nrow(caro), by = 3)
caro_3 <- caro %>% slice(seq_3)

seq_6 <- seq(from = 1, to = nrow(caro), by = 6)
caro_6 <- caro %>% slice(seq_6)

seq_9 <- seq(from = 1, to = nrow(caro), by = 9)
caro_9 <- caro %>% slice(seq_9)
```

```{r}
#Validation:
nrow(caro)
## [1] 200
nrow(caro_3)
## [1] 67
nrow(caro_6)
## [1] 34
nrow(caro_9)
## [1] 23
```

Now calculate timelag, steplength and speed:  
```{r}
#check if different individuals:
caro %>% group_by(TierName) %>% summarise(n_observations=n())
```
-> only Caro.  

```{r}
#1 minute timelag
#timelag
caro <- mutate(caro, timelag = as.numeric(difftime(lead(DatetimeUTC), DatetimeUTC)))

#steplength
caro <- caro %>% mutate(steplength_m = sqrt((E-lead(E))^2 + (N-lead(N))^2))

#speed
caro <- caro %>% mutate(speed_ms = steplength_m/timelag)
```

```{r}
#3 minutes timelag
#timelag
caro_3 <- mutate(caro_3, timelag = as.numeric(difftime(lead(DatetimeUTC), DatetimeUTC)))

#steplength
caro_3 <- caro_3 %>% mutate(steplength_m = sqrt((E-lead(E))^2 + (N-lead(N))^2))

#speed
caro_3 <- caro_3 %>% mutate(speed_ms = steplength_m/timelag)
```

```{r}
#6 minutes timelag
#timelag
caro_6 <- mutate(caro_6, timelag = as.numeric(difftime(lead(DatetimeUTC), DatetimeUTC)))

#steplength
caro_6 <- caro_6 %>% mutate(steplength_m = sqrt((E-lead(E))^2 + (N-lead(N))^2))

#speed
caro_6 <- caro_6 %>% mutate(speed_ms = steplength_m/timelag)
```

```{r}
#9 minutes timelag
#timelag
caro_9 <- mutate(caro_9, timelag = as.numeric(difftime(lead(DatetimeUTC), DatetimeUTC)))

#steplength
caro_9 <- caro_9 %>% mutate(steplength_m = sqrt((E-lead(E))^2 + (N-lead(N))^2))

#speed
caro_9 <- caro_9 %>% mutate(speed_ms = steplength_m/timelag)
```

Plot trajectories:  

```{r}
ggplot() +
  geom_path(data = caro_3, aes(x = E, y = N, colour="3 minutes")) +
  geom_point(data = caro_3, aes(x = E, y = N, colour="3 minutes"), size=2) +
  geom_path(data = caro, aes(x = E, y = N, colour = '1 minute'), alpha = 1/2) +
  geom_point(data = caro, aes(x = E, y = N, colour="1 minute"), size=1, alpha = 1/5) +
  ggtitle("Comparing original with 3-minutes resampled data") +
  labs(colour = "Trajectory")
```

```{r}
ggplot() +
  geom_path(data = caro_6, aes(x = E, y = N, colour="6 minutes")) +
  geom_point(data = caro_6, aes(x = E, y = N, colour="6 minutes"), size=2) +
  geom_path(data = caro, aes(x = E, y = N, colour = '1 minute'), alpha = 1/2) +
  geom_point(data = caro, aes(x = E, y = N, colour="1 minute"), size=1, alpha = 1/5) +
  ggtitle("Comparing original with 6-minutes resampled data") +
  labs(colour = "Trajectory")
```

```{r}
ggplot() +
  geom_path(data = caro_9, aes(x = E, y = N, colour="9 minutes")) +
  geom_point(data = caro_9, aes(x = E, y = N, colour="9 minutes"), size=2) +
  geom_path(data = caro, aes(x = E, y = N, colour = '1 minute'), alpha = 1/2) +
  geom_point(data = caro, aes(x = E, y = N, colour="1 minute"), size=1, alpha = 1/5) +
  ggtitle("Comparing original with 9-minutes resampled data") +
  labs(colour = "Trajectory")
```

Compare the speeds in a line plot:  
```{r}
ggplot() +
  geom_path(data = caro_9, aes(x = DatetimeUTC, y = speed_ms, colour="9 minutes")) +
  geom_path(data = caro_6, aes(x = DatetimeUTC, y = speed_ms, colour="6 minutes")) +
  geom_path(data = caro_3, aes(x = DatetimeUTC, y = speed_ms, colour="3 minutes")) +
  geom_path(data = caro, aes(x = DatetimeUTC, y = speed_ms, colour = '1 minute')) +
  
  ggtitle("Comparing speeds") +
  labs(colour = "Speed m/s")
```
## Task 6: Add your movement data to your repository


## Task 7: Explore your movement data


